AppPAL for Android

### 2016-01-08 paper.tex
copyright and hanging package
last minute tweaks

\usepackage{graphicx}
\usepackage{defaultstyle}
\usepackage{localstyle}

@@

\begin{abstract}
  It can be difficult to find mobile apps that respect your security and privacy.
  Businesses rely on employees enforcing company mobile device policies correctly.
  Users must judge apps by the information shown to them by the store.
  Only 17\% of users pay attention to an apps permissions during installation~\cite{Felt:2012hm} and most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
  To address these problems, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2009ko}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
  AppPAL also acts as a ``glue'', allowing connection to a variety of local constraint checkers (e.g., static analysis tools, packager manager checks) to combine their results.
  As well as introducing AppPAL and some examples, we apply it to explore whether real users follow certain intended policies in practice, finding privacy preferences and actual behaviour are not always aligned, in the absence of a rigorous enforcement mechanism.
  \end{abstract}

@@

App stores give some information about their apps; descriptions, screenshots and review scores.
Android apps show a list of permissions when they're first installed.
In Android Marshmallow apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision which apps to use and which permissions to grant must be made by the device user.

Some apps are highly undesirable.
Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work.
An employer may restrict which apps their employees can use.
The IT department may set a policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
Some users worry apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
Such a user avoids apps which can access their location, or address book.
They may apply their own personal security policies when downloading and running apps.

These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
An alternative is to write the policy down and make the computer enforce it.
To implement this we use a logic of authorization---a language designed to express rules about permissible actions.
The policy is written in the logic and enforced by checking the policy is satisfied.

We present AppPAL, an instantiation of Becker~\etal's SecPAL~\cite{Becker:2006vh} with constraints\footnote{A constraint makes use of information checkable using information external to the language, such as the time of day or output of a static analysis tool.} and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
AppPAL allows us to enforce the policies on a device.
We can express trust relationships amongst these parties and use constraints to do additional checks, such as using static analysis results.
This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.

A user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
Using AppPAL we can write policies for work and home, and decide which policy to enforce using a user's location, or the time of day:

\noindent
\begin{minipage}{\linewidth}
  \begin{minipage}{0.5\linewidth}
    \begin{lstlisting}
'alice' says App isRunnable
  if 'home-policy' isMetBy(App)
  where at('work') = false.
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \begin{lstlisting}
'alice' says App isRunnable
  if 'work-policy' isMetBy(App)
  where beforeHourOfDay('17') = true.
    \end{lstlisting}
  \end{minipage}
\end{minipage}

We can delegate policy specification to third parties or roles, and assign principals to roles:
\begin{lstlisting}
'alice' says 'it-department' can-say 'work-policy' isMetBy(App).
'alice' says 'alice' can-act-as 'it-department'.
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization.
For example it would be okay allowing a photography app access to the camera, but not to location data if the user doesn't want their photos geotagged.
\begin{lstlisting}
'alice' says App isRunnable
  if 'permissions-policy' isMetBy(App).
'alice' says 'permissions-policy' isMetBy(App)
  if App isAnApp
  where
    category(App, 'Photography'),
    hasPermission(App, 'LOCATION') = false,
    hasPermission(App, 'CAMERA') = true.
\end{lstlisting}

There has been a great amount of work on developing app analysis tools for Android.
Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Arzt:2014kf,Li:2015wo} can do taint and control flow analysis; sometimes even between app components.
Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between entire apps.
ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
MalloDroid~\citep{Fahl:2012dj} looks for apps configured to use SSL incorrectly (for instance by not verifying hostnames or certificates).
Many others exist checking and certifying other aspects of app behaviour.

AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owners are trying to enforce.
This avoids creating tools with hard-coded fixed policies.
For example a store might not want to sell apps with SSL errors or malicious apps as determined by their anti-virus.
Using AppPAL we can combine tools for checking apps to implement the store's policies.
\begin{lstlisting}
'play-store' says App isSellable
  if App isAnApp
  where mallodroidCheck(App) = true,
         mcafeeAVCheck(App) = true.
\end{lstlisting}

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy.
Alternately Emma's policy could be partially enforced using existing tools.
\emph{Google's Device Policy for Android}~\cite{GoogleAppsDeviceP:tz} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}.

We could implement Emma's policy using existing tools (such as an AV checker, and a taint analysis tool like Flowdroid~\cite{Arzt:2014kf,Li:2015wo}) but it is a clumsy solution---they are not flexible.
Each has to be configured separately to implement only part of the policy.
If Emma changes her policy or Alice changes jobs she needs to recheck her apps and then alter or remove the software on her phone to ensure compliance.
It isn't clear what an app must do to be run, or what checks have been done if it is already running on the phone.
The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.

What happens when Alice goes home?
Emma shouldn't be able to overly control what Alice does in her private life.
Alice might not be allowed to use location tracking apps at work but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disable at run time.
Can we enforce different policies at different times or locations?

@@

We propose a change to the mobile ecosystem shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the explicitly specified store policies.
Developers could decide which stores to sell their apps in on the basis of policies about stores.

@@

In \autoref{sec:problem} Alice and Emma had policies they wanted to enforce but no means to do so.
Instead of using several tools to enforce Emma's policy disjointedly, we could use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.

SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
We present AppPAL as a modified form of SecPAL, targeting apps on mobile devices.

\begin{figure}
  \begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
'alice' says 'emma' can-say inf
  App isRunnable. @\label{lst:corporate_1}@

@@

'emma' says 'google-play' can-say
  App isBuyable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.5\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15,
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
'emma' says 'anti-virus-policy' isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcafeeAVCheck(App) = true.

@@

In \autoref{lst:corporate_1} Alice less Emma  specify whether an \code{App} (a variable) \code{isRunnable};
she allows her to delegate the decision further (\code{can-say}\/~\code{inf}).
Emma specifies her concerns as policies to be met in \autoref{lst:corporate_2}:
  if Emma is convinced all these are met then she will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
  with \code{"google-play"} as the decider of what is buyable or not.
Google is not allowed to delegate the decision further, i.e.~Google is not allowed to specify Amazon as a supplier of apps as well.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5} using a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
She needs to collect statements to show the app meets the \code{isRunnable} predicate.
Namely:
\begin{itemize}
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the second statement AppPAL must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the third statement the AppPAL checker needs to examine the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if it is running on a device.

@@

In SecPAL the precise nature of predicates and constraints is left open.
In instantiating SecPAL, AppPAL makes the predicates and constraints explicit.
AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
Additional predicates can be created in the policy files, however constraints are implemented individually.
For example on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.

\begin{table}
  \footnotesize
  \newcommand{\style}[1]{{\footnotesize #1}}
  \begin{tabulary}{\linewidth}{l L}
  \toprule
  \style{Name}                                             & \style{Description}                                                                                                                        \\
  \midrule
  \style{App \texttt{isRunnable}}                          & \style{Says an app can be run.}                                                                                                            \\
  \style{App \texttt{isInstallable}                      } & \style{Says an app can be installed.                                                                                                     } \\
  \style{App \texttt{isAnApp}                            } & \style{Tells AppPAL that an app exists.                                                                                                  } \\
  \style{Policy \texttt{isMetBy(}App\texttt{)}           } & \style{Used to split policies into smaller components.}                                                                                    \\
  \style{\texttt{hasPermission(}App, Permission\texttt{)}} & \style{Constraint to check if an app has a permission.                                                                                   } \\
  \style{\texttt{beforeHourOfDay(}time\texttt{)}         } & \style{Constraint used to check the time.                                                                                                } \\
  \style{Tool\texttt{Check(}App, Property)               } & \style{Constraint to run an analysis tool on an app.                                                                        }              \\
  \bottomrule
\end{tabulary}
\caption{AppPAL predicates and constraints.}
\label{tab:predicates}
\end{table}

Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one fails.
Some static analysis tools are not quick to run.
Even taking minutes to run a battery draining analysis can be undesirable:
if a user wants to download an app quickly they may not be willing to wait to check that a policy is met.

In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
In Emma's policy we checked this using the app's permissions;
if the app couldn't get access to the GPS sensors (using the permissions) then it met this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Arzt:2014kf,Li:2015wo}).
Our policy now becomes:

@@

On a mobile device memory is at a premium.
We would like to keep the assertion context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
When delegating we will also be able to request facts from the delegated party dynamically; though implementing this is future work.

@@

When AppPAL runs on a mobile phone apps should be checked as they are installed.
Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
Downloading and installing an app takes about 30 seconds on a typical Android phone over wifi.
If checking a policy delays this even further a user may become annoyed and disable AppPAL.

The policy checking procedure is at its slowest when having to delegate repeatedly;
   the depth of the delegation tree is the biggest factor for slowing the search.
Synthetic benchmarks were created to check that the checking procedure performed acceptably.
Each benchmark consisted of a chain of delegations.
The \emph{1 to 1} benchmark consists of a repeated delegation between all the principals.
In the \emph{1 to 2} benchmark each principal delegated to 2 others and in the \emph{1 to 3} benchmark each principal delegated to 3 others.
These benchmarks are reasonable as they model the slowest kinds of policies to evaluate---though worse ones could be designed by delegating even more.

@@

Throughout we have asserted that users have policies and that there is a need for policy enforcement tools.
Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq} demonstrating a refinement of Westin's privacy segmentation index~\cite{Krane:2002jo}.
Using app installation data from Carat~\cite{Oliner:2013ht,Chia:2012gz} we used AppPAL to find the apps satisfying each policy Lin~\etal~identify and measure the extent that each user was following a policy.

Lin~\etal~identified four types of user.
The \emph{Conservative} (C) users were uncomfortable allowing an app access to any personal data for any reason.
The \emph{Unconcerned} (U) users felt okay allowing access to most data for almost any reason.
The \emph{Advanced} (A) users were comfortable allowing apps access to location data but not if it was for advertising reasons.
Opinions in the largest cluster, \emph{Fencesitters} (F), varied but were broadly against collection of personal data for advertising purposes.
We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
These simplify the privacy policies identified by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission.
Using AppPAL we could write more precise rules if we could determine why each permission was requested.
Lin~\etal~used Androguard~\cite{Desnos:ub} as well as manual analysis to determine the precise reasons for each permission~\cite{Sadeh:2014vq}.

>>>>

\usepackage{graphicx}
\usepackage{defaultstyle}
\usepackage{localstyle}
\usepackage{hanging}

@@

\begin{abstract}
  It can be difficult to find mobile apps that respect one's security and privacy.
  Businesses rely on employees enforcing company mobile device policies correctly.
  Users must judge apps by the information shown to them by the store.
  Studies have found that most users do not pay attention to an apps permissions during installation~\cite{Felt:2012hm} and most users do not understand how permissions relate to the capabilities of an app~\cite{Kelley:2012bw}.
  To address these problems and more, we present AppPAL: a machine-readable policy language for Android that describes precisely when apps are acceptable.
  AppPAL goes beyond existing policy enforcement tools, like Kirin~\cite{Enck:2009ko}, adding delegation relationships to allow a variety of authorities to contribute to a decision.
  AppPAL also acts as a ``glue'', allowing connection to a variety of local constraint checkers (e.g., static analysis tools, packager manager checks) to combine their results.
  As well as introducing AppPAL and some examples, we apply it to explore whether users follow certain intended policies in practice, finding privacy preferences and actual behaviour are not always aligned in the absence of a rigorous enforcement mechanism.
  \end{abstract}

@@

App stores give some information about their apps; descriptions, screenshots and review scores.
Android apps show a list of permissions when they're first installed.
In Android Marshmallow apps will display permissions requests when the app first tries to access sensitive data (such as contacts or location information).
Users do not understand how permissions relate to their device~\cite{Felt:2012hm,Thompson:2013eb}.
Ultimately the decision of which apps to use and which permissions to grant must be made by the device user.

Some apps are highly undesirable.
Many \ac{pup} are being propagated for Android devices~\cite{Truong:2014bi,Svajcer:2013tp}.
Employees are increasingly using their own phones for work.
An employer may restrict which apps their employees can use.
The IT department may set a mobile device policy---a series of rules describing what kinds of apps may be used and how---to prevent information leaks.
Some users worry that apps will misuse their personal data---sending their address book or location to an advertiser without their permission.
Such a user avoids apps which can access their location, or address book; they may apply their own personal security policies when downloading and running apps.

These policies can only be enforced by the users continuously making the correct decision when prompted about apps.
An alternative is to write the policy down and make the computer enforce it.
To implement this we propose using a logic of authorization---a language designed to express rules about permissible actions.

We present AppPAL, an instantiation of Becker~\etal's
SecPAL~\cite{Becker:2006vh} with constraints (statements checkable using information external to the language such as the time of day or static analysis tools) and predicates that allow us to decide which apps to run or install.
The language allows us to reason about apps using statements from third parties.
AppPAL allows us to enforce the policies on a device.
We can express trust relationships amongst these parties and use constraints to do additional checks, such as using security checks.
This lets us enforce more complex policies than existing tools such as Kirin~\cite{Enck:2009ko} which are limited to permissions checks.
Policies can be enforced by the stores selling the apps, on the devices installing apps or by third-parties providing app vetting services.

Consider the following example:
  a user, Alice, may have rules she has to follow when using apps for work and her own policies when using apps at home in her private life.
Using AppPAL we can write policies for work and home, and decide which policy to enforce using a user's location, or the time of day:

\noindent
\begin{minipage}{\linewidth}
  \begin{minipage}{0.5\linewidth}
    \begin{lstlisting}
'alice' says App isRunnable
  if 'home-policy' isMetBy(App)
  where at('work') = false.
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \begin{lstlisting}
'alice' says App isRunnable
  if 'work-policy' isMetBy(App)
  where beforeHourOfDay('17') = true.
    \end{lstlisting}
  \end{minipage}
\end{minipage}
\noindent We can delegate policy specification to third parties or roles, and assign principals to roles:
\begin{lstlisting}
'alice' says 'it-department' can-say 'work-policy' isMetBy(App).
'alice' says 'alice' can-act-as 'it-department'.
\end{lstlisting}
We can write policies specifying which permissions an app must or must not have by its app store categorization.
For example, it would be okay allowing a photography app access to the camera, but not to allow access to location data if the user doesn't want their photos geotagged.
\begin{lstlisting}
'alice' says App isRunnable
  if 'permissions-policy' isMetBy(App).
'alice' says 'permissions-policy' isMetBy(App)
  if App isAnApp
  where
    category(App, 'Photography'),
    hasPermission(App, 'LOCATION') = false,
    hasPermission(App, 'CAMERA') = true.
\end{lstlisting}

There has been a great amount of work on developing app analysis tools for Android.
Tools such as Stowaway~\cite{Felt:2011kj} detect over-privileged apps.
TaintDroid~\cite{Enck:2010uw} and FlowDroid~\cite{Arzt:2014kf,Li:2015wo} can do taint and control flow analysis; sometimes even between app components.
Other tools like QUIRE~\cite{Bugiel:2012ui} can find privilege escalation attacks between apps.
ScanDAL~\cite{Kim:2012vt} and SCanDroid~\cite{Fuchs:2009vi} help detect privacy leaks.
Appscopy~\cite{Feng:kPGZr_ja} searches for specific kinds of malware.
Tools like DroidRanger~\cite{Zhou:2012tb} scan app markets for malicious apps.
Various tools such as AppGuard~\cite{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\cite{Jeon:2012ki} or AppFence~\cite{Hornyack:2011wq} can control the permissions or data an app can get.
MalloDroid~\citep{Fahl:2012dj} looks for apps configured to use SSL incorrectly (for instance by not verifying hostnames or certificates).

AppPAL can act as a \emph{``glue"} between static analysis tools and the app installation policies device owners are trying to enforce.
This avoids creating tools with hard-coded fixed policies.
 For example a store might not want to sell apps with SSL errors or apps flagged by an anti-virus tool.
Using AppPAL we can combine tools for checking apps to implement the store's policies.
\begin{lstlisting}
'play-store' says App isSellable
  if App isAnApp
  where mallodroidCheck(App) = true,
         mcafeeAVCheck(App) = true.
\end{lstlisting}
No additional attempt is made to ensure these static analysis tools are sound.
The policy designer must be aware of the tool's limitations.
Black or whitelisting may have to be used to avoid some false positives or negatives.

@@

To ensure this policy is met Alice promises to follow it.
She might even sign a document promising never to break the rules within the policy.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
Alternatively Emma's policy could be partially enforced using existing tools.
\emph{Google's Device Policy for Android}~\cite{GoogleAppsDeviceP:tz} could configure Alice's device to disallow apps from outside the Google Play Store and let Emma set the permissions granted to each app~\cite{AndroidMPermission:2015uq}.

We could implement Emma's policy using existing tools (such as an AV checker, and a taint analysis tool like Flowdroid~\cite{Arzt:2014kf,Li:2015wo}) but it is a clumsy solution---they are not flexible.
Each has to be configured separately to implement only part of the policy.
If Emma changes her policy or Alice changes jobs she must recheck her apps and then alter or remove the software on her phone to ensure compliance.
It isn't clear what an app must do to be run, or what checks have been done if it is already running on the phone.
The relationship between Alice (the user), Emma (the policy setter) and the tools Emma trusts to implement her policy isn't immediately apparent.

What happens when Alice goes home?
Emma shouldn't be able to overly control what Alice does in her private life.
Alice might not be allowed to use location tracking apps at work but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
Some mobile OSs, such as iOS and the latest version of Android, allow app permissions to be enabled and disabled at run time.
Can we enforce different policies at different times or locations?

@@

We propose using the mobile ecosystem shown in \autoref{fig:ecosystem}.
People have policies which are enforced by AppPAL on their devices.
They can be composed with policies from employers or others to create enhanced devices that ensure apps meet the policies of their owners.
The device can make use of vetting services which run tools to infer complex properties about apps.
Users can buy from enhanced stores which ensure the only apps they sell are the apps which meet the store's explicit policies, or ones requested by users.
Developers could decide which stores to sell their apps in on the basis of policies about stores.

@@

In \autoref{sec:problem}, Alice and Emma had policies they wanted to enforce but no means to do so.
Instead of using several tools to enforce Emma's policy disjointedly, we could use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.

SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\cite{Becker:2009ula} and data-sharing~\cite{Aziz:2011vt}.
We present AppPAL as a modified form of SecPAL, aimed at mobile apps.

Other access control languages, such as XACML~\cite{Oasis:1M71_oFS}, could also have been used as the basis for AppPAL.
We chose to use SecPAL as the basis, however, as it was more lightweight, designed to be readable, and simpler to extend with additional constraints than XACML.

\begin{figure}
  \begin{minipage}[t]{0.44\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
'alice' says 'emma' can-say inf
  App isRunnable. @\label{lst:corporate_1}@

@@

'emma' says 'google-play' can-say
  App isBuyable. @\label{lst:corporate_4}@
    \end{lstlisting}
  \end{minipage}\begin{minipage}[t]{0.56\textwidth}
    \begin{lstlisting}[numbers=left,
                       escapeinside={@}{@},
                       firstnumber=15,
                       basicstyle=\scriptsize\ttfamily,
                       stringstyle=\scriptsize\sffamily,
                       keywordstyle=\scriptsize\slshape]
'emma' says 'anti-virus-policy' isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcafeeAVCheck(App) = true.

@@

In \autoref{lst:corporate_1} Alice lets Emma specify whether an \code{App} (a variable) \code{isRunnable};
she allows her to delegate the decision (\code{can-say}\/~\code{inf}).
Emma specifies her concerns as policies to be met in \autoref{lst:corporate_2}:
  if Emma is convinced that these are met then she will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-policy} if the \code{App isBuyable};
  with \code{"google-play"} deciding of what is buyable or not.
Google is not allowed to delegate the decision further, i.e.~Google is not allowed to specify Amazon as a supplier of apps as well.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5} using a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{'no-location-permissions'} rules implement the \code{'no-tracking-policy'} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
She collects statements to show the app meets the \code{isRunnable} predicate.
\begin{itemize}
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma that the app came from a reputable source.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if the app has no location permissions.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the second statement AppPAL must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as its signatures are updated;
  so the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the third statement the AppPAL checker needs to examine the permissions of the app.
It could do this by looking in the \texttt{MANIFEST.xml} inside the app itself, or through the Android package manager if it is running on a device.

@@

In SecPAL the precise nature of predicates and constraints is left open.
In instantiating SecPAL, AppPAL makes the predicates and constraints explicit.
AppPAL policies can make use of the predicates and constraints in \autoref{tab:predicates}.
Additional predicates can be created in the policy files, however constraints must be implemented individually.
For example, on Android the \texttt{hasPermission} constraint uses the Android package manager to check what permissions an app requests, but the Java version uses the Android platform tools to check.

\begin{table}
  \footnotesize
  \newcommand{\style}[1]{{\footnotesize #1}}
  \begin{tabulary}{\linewidth}{l L}
  \toprule
  \style{Name}                                             & \style{Description}                                                                                                                        \\
  \midrule
  \style{App \texttt{isRunnable}}                          & \style{Says an app can be run.}                                                                                                            \\
  \style{App \texttt{isInstallable}                      } & \style{Says an app can be installed.                                                                                                     } \\
  \style{App \texttt{isAnApp}                            } & \style{Tells AppPAL that an app exists.                                                                                                  } \\
  \style{Policy \texttt{isMetBy(}App\texttt{)}           } & \style{Used to split policies into smaller components.}                                                                                    \\
  \style{\texttt{hasPermission(}App, Permission\texttt{)}} & \style{Constraint to check if an app has a permission.                                                                                   } \\
  \style{\texttt{beforeHourOfDay(}time\texttt{)}         } & \style{Constraint used to check the time.                                                                                                } \\
  \style{Tool\texttt{Check(}App, Property)               } & \style{Constraint to run an analysis tool on an app.                                                                        }              \\
  \bottomrule \\
\end{tabulary}
\caption{AppPAL predicates and constraints.}
\label{tab:predicates}
\end{table}

Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one fails.
Some static analysis tools are not quick to run.
Even taking minutes to run a battery draining analysis can be undesirable:
if a user wants to download an app quickly they may not be willing to wait to check that a policy is met.
In that case, it may be preferable to delegate to an online database.

In \autoref{sec:problem} and \autoref{sec:idea} we described a \emph{no-tracking-policy} to prevent a user's location being leaked.
In Emma's policy we checked this using the app's permissions;
if the app couldn't get access to the GPS sensors (using the permissions) then it meets this policy.
Some apps may want to access this data, but may not leak it.
We could use a taint analysis tool to detect this (e.g.~FlowDroid~\cite{Arzt:2014kf,Li:2015wo}).
Our policy becomes:

@@

On a mobile device memory is at a premium.
We want to keep the assertion context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.
This gives us greater control of the evaluation and how the assertion context is created.
For example, when checking the \code{isAnApp} predicate;
  we can fetch the assertion that the subject is an app based on the app in question.
When delegating we will also be able to request facts from the delegated party dynamically (although this is not yet implemented).

@@

When AppPAL runs on a mobile phone, apps should be checked as they are installed.
Since policy checks may involve inspecting many rules and constraints one may ask whether the checking will be acceptably fast.
Downloading and installing an app takes about 30 seconds on a typical Android phone over wifi.
If checking a policy delays this even further a user may become annoyed and disable AppPAL.

The policy checking procedure is at its slowest when having to delegate repeatedly;
   the depth of the delegation tree is the biggest factor for slowing the search.
Synthetic benchmarks were created to check that the checking procedure performed acceptably.
Each benchmark consisted of a chain of delegations.
The \emph{1 to 1} benchmark consists of a repeated delegation between all the principals.
In the \emph{1 to 2} benchmark each principal delegated to 2 others and in the \emph{1 to 3} benchmark each principal delegated to 3 others.
These benchmarks are reasonable as they model the slowest kinds of policies to
evaluate---though worse ones could be designed by delegating even more or triggering an expensive constraint check.

@@

Throughout we have asserted that users often have informal policies and that there is a need for policy enforcement tools.
Corporate mobile security \ac{byod} policies have started appearing and NIST have issued recommendations for writing them~\cite{Scarfone:2009vy,Souppaya:2013jf}.
In a study of 725 Android users, Lin~\etal~found four patterns that characterise user privacy preferences for apps~\cite{Sadeh:2014vq} demonstrating a refinement of Westin's privacy segmentation index~\cite{Krane:2002jo}.
Using app installation data from Carat~\cite{Oliner:2013ht,Chia:2012gz} we used AppPAL to find the apps satisfying each policy Lin~\etal~identify and measure the extent that each user was following a policy.

Lin~\etal~identified four types of user.
The \emph{Conservative} (C) users were uncomfortable allowing an app access to any personal data for any reason.
The \emph{Unconcerned} (U) users felt okay allowing access to most data for almost any reason.
The \emph{Advanced} (A) users were comfortable allowing apps access to location data but not if it was for advertising.
Opinions in the largest cluster, \emph{Fencesitters} (F), varied but were broadly against collection of personal data for advertising.
We wrote AppPAL policies to describe each of these behaviours as increasing sets of permissions.
These simplify the privacy policies identified by Lin~\etal~as we do not take into account the reason each app might have been collecting each permission
(we could write more precise rules if we could determine why each permission was requested).
Lin~\etal~used Androguard~\cite{Desnos:ub} as well as manual analysis to determine the precise reasons for each permission~\cite{Sadeh:2014vq}.

\newcommand{\tabtitle}[1]{\textbf{\footnotesize #1}}
\begin{center}
  \begin{tabular}{ r l l l l }
%                                     & \rothead{Conservative} & \rothead{Advanced} & \rothead{Fencesitter} & \rothead{Unconcerned} \\
    \toprule
    \tabtitle{Policy}                 & \tabtitle{C}           & \tabtitle{A}       & \tabtitle{F}          & \tabtitle{U}          \\
    \midrule
    \texttt{GET\_ACCOUNTS}            & \xmark                 & \xmark             & \xmark                & \xmark                \\
    \texttt{ACCESS\_FINE\_LOCATION}   & \xmark                 & \xmark             & \xmark                &                       \\
    \texttt{READ\_CONTACT}            & \xmark                 & \xmark             & \xmark                &                       \\
    \texttt{READ\_PHONE\_STATE}       & \xmark                 & \xmark             &                       &                       \\
    \texttt{SEND\_SMS}                & \xmark                 & \xmark             &                       &                       \\
    \texttt{ACCESS\_COARSE\_LOCATION} & \xmark                 &                    &                       &                       \\
    \bottomrule
  \end{tabular}
\end{center}
